### 2018-11-15 design.tex
reorg

\section{SPP-Perc}

\subsection{Changes made to original SPP}
To modify the SPP Design to suit our scheme, the following changes were made:

\begin{itemize}
\item \textbf{Original Thresholds discarded}\newline In the original
  SPP design, an internal confidence mechanism was developed which was
  compared against a preset PF\_THRESHOLD (T\textsubscript{p}) to
  decide whether to accept or reject the prefetch suggestion.
  Furthermore, the accepted suggestions were thresholded against FILL
  \_THRESHOLD (T\textsubscript{f}) to decide the fill level: L2 Cache
  vs Last Level Cache.  T\textsubscript{p} and T\textsubscript{f} were
  set to 25 and 90 respectively, on the scale of 0 to 100.  In
  SPP-Perc, the perceptron sum is used to make these decisions.  Hence
  these two thresholds are no longer needed

\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline
  Original SPP would not consider the slots available in the L2 MSHR
  queue.  It had an assumption that at no time can the number of
  prefetches suggested by SPP could exceed the capacity of the queue.
  While it was not proven in the paper, the internal confidence
  mechanism along with the thresholds made sure that the above
  assumption was maintained.  In SPP-Perc since there can be no
  assumption on the degree of aggressiveness - because SPP is being
  tuned to its lowest thresholds, we explicitly check that at no point
  should the number of suggested prefetches exceed L2 MSHR queue.

\item \textbf{Enhanced Filter}\newline Original SPP introduces a
  concept of Prefetch Filter, which keeps a track of up to 1024 L2
  Cache prefetches suggested by SPP.  The Filter was modified to store
  meta-data like program counter, memory access address etc.  This is
  needed to train the perceptron when the result of that particular
  prefetch gets available.  At training time, this data is used to
  retrieve the state of the prefetcher when the prediction was made.

\item \textbf{Reject Filter introduced}\newline The Prefetch Filter
  mentioned above keeps a track of the prefetches that actually
  happened in the L2 Cache.  Correspondingly, we introduced a reject
  filter.  It keeps a track of the prefetches that were suggested by
  the base SPP engine but were rejected by the perceptron.  In case it
  is detected that a future demand fetch could have used this rejected
  prefetch, the perceptron weights are updated to reflect the same

\end{itemize}

\subsection{The Perceptron Layer}
As discussed earlier, the perceptron layer was introduced to act as a
throttler to the underlying SPP engine.  This section explores the
details of the perceptron layer and the range of features that are
ultimately used by the perceptron for correlating the prefetches.

The neural network weights are stored in separate tables for each
feature.  The indexing of each feature is done using different number
of bits.  At most 12 bits of a feature are used to index into the
perceptron table.  Certain features require more resolution power
\textit{i.e.}, full 12 bits of indexing.  On the other hand, some
features requiring lower resolution required as low as 7 bits of
indexing.  The variable indexing was determined by studying the
features and fine-tuned empirically so as to achieve a good accuracy
vs area trade-off.  Exact details are mentioned under `Area Overheads'
in Section XX.

A single entry in the table corresponds to a perceptron weight.  Each
weight is a 5 bit counter - saturating at -16 and +15.  Each feature
gets its own dedicated table.  Our proposed design uses 9 features.
As the program execution begins, all the weights are initialized to 0.

\textbf{INFERENCING}\newline In the Champsim simulator, the
prefetching method is triggered on every L2 Cache demand access.  At
that point, the prefetcher has the option to do or not do prefetch -
in case it does, then how many cache lines to prefetch.  The suggested
prefetches can be places either in L2 Cache or in Last Level Cache.
The prefetcher gets to decide this, and is usually done based on an
internal confidence mechanism.

When SPP gets triggered, it suggests up to 4 (=Associativity of the
Pattern Table) candidates for prefetch based on the demand access
address.  In each subsequent lookahead depth, SPP suggests up to 4
more prefetch candidates, as explained in section\_\_.  All these
recommended prefetches are tested in the perceptron throttler before
they can access the Prefetch Filter, as in the original SPP data-path.

The perceptron looks at the micro-architectural state, or `features',
at that instant.  Each feature indexes into the hash of up to
4096-entry table dedicated for that feature.  Feature \#1 indexes into
a particular entry in table \#1 and so on.

Once all the weights are retrieved, they are added together, much like
the traditional perceptron dot product.  On obtaining the overall sum,
it is thresholded based on the preset threshold PERC\_THRESHOLD\_LO.
Only the prefetch candidates with the perceptron sum higher that the
threshold qualify for prefetching.  The prefetches that qualify the
perceptron throttler stage go through the Prefetch Filter stage, just
as in the SPP design.  The Prefetch Filter eliminates any redundant
prefetch suggestions.  In SPP-Perc, all the meta-data that is required
to recreate the features is also stored in the Prefetch Filter.  At a
later stage, when the feedback of the current prefetched line is
available, the stored data can be used to train the perceptron.

The above generated perceptron sum is also used to decide L2 Cache vs
Last Level Cache placement of the prefetch cache line.  All prefetch
candidates qualified till this stage are thresholded against
PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.

The perceptron sum can also be seen as the sum of the individual
contribution per feature.  The value of each contribution corresponds
to the extent of confidence of the final output with that given
feature.  By summing the individual contributions, the final
perceptron sum denotes the overall confidence for that prefetch
suggestion.  In a way, by thresholding the perceptron sum on two
different thresholds, we are diving the confidence scale of the
prefetch suggestion into three bins.  The first bin corresponding to
the lowest confidence leads to prefetch candidate being rejected.  The
next bin corresponds to prefetches with a moderate confidence level.
Such prefetches are directed to fill the bigger Last Level Cache and
potentially not pollute the more scarce L2 Cache.  The highest
confidence prefetches are filled in the L2 Cache.

In addition to the Prefetch Filter mentioned above, SPP-Perc also
maintains a `Reject Filter'.  Reject filter is a 1024-entry deep 1-way
associative table.  In case that a prefetch suggestion is rejected by
perceptron throttler, it instead gets logged into the Reject Filter.
The Reject Filter keeps track of 1024 entries that were suggested by
the Pattern Table but got rejected by the perceptron throttler.  This
can be used to train the perceptron to avoid false negatives
\textit{i.e.}, cases where prediction was to reject the prefetch but
the prefetch would have been useful.

As previously stated, an enhancement to the original SPP design
includes storing all the meta-data describing the program state at the
instant of prefetching, in the Prefetch Filter.  This information is
useful when the perceptron needs to be updated subsequently.

\textbf{TRAINING}\newline In the prefetching environment, feedback of
a prefetch is received whenever there is an eviction or a demand
access from L2 Cache.  This acts as a trigger to update all the
baseline SPP entries and to also train the perceptron throttler.
Training of the perceptron throttler involves restoring the meta-data
that was stored in Prefetch Filter or the Reject Filter.  This is done
by indexing into the corresponding filters using the cache line
address of the block being used to train. (10-bits indexing + 6-bits
tag matching).  Once the state of the program at the time of
prefetching is available, it is used to index into the perceptron
weights table.

If a demand access block which triggers the training was tagged as a
valid prefetch in the Prefetch Filter, it means that the earlier
prefetch prediction was correct.  In that case the perceptron weights
are incremented by 1 if the predicted sum does not cross a pre-defined
threshold.  In case it is a cache block eviction that led to training
and the corresponding valid entry was found in the Prefetch Filter, it
shows that the prediction made by the perceptron was wrong.
Perceptron should have ideally rejected the prefetch suggestion as a
low-confidence prefetch.  Here the weights are decremented by 1 to
reflect the same.  In either case, weights are saturated at -16 or
+15.

A secondary training mechanism also kicks in during demand fetches.
Before the demand access triggers the next set of prefetches, it
checks the Reject Filter for the valid entry.  In case it is a hit,
that means that the corresponding cache line was initially suggested
by the underlying SPP engine but rejected by the perceptron throttler.
This means that perceptron should have been more confident about that
particular prefetch.  Once such a scenario is identified, the state of
the execution at the time of prefetch is retrieved from the Reject
Table.  The retrieved data is used to index into the various weights
tables of the perceptron and the corresponding values are updated by
+1, saturating between and -16 and +15.  This update reflects
increased confidence for the prediction corresponding to that
prefetch.

This mechanism allows us to solve the classical problem of exploiting
`lost opportunity'.  In prior perceptron based implementations (and in
general, prefetching algorithms), there is usually no way of knowing
the result of not prefetch a particular line.  Our two-step SPP-Perc
allows us to overcome that issue in a simple manner.

\subsection{Features used by Perceptron}

Here we discuss the various features which help correlate the
prefetching decision with the program behaviour
\begin{itemize}
\item \textbf{Base address} of the demand access which triggered the
  prefetch.  Since prefetch are triggered from the L2 demand access,
  they tend to be correlated with the triggering address.

\item \textbf{Cache line} and \textbf{Page address}: These two
  separate features are derived from the base address which triggered
  the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
  base\_addr >> LOG2\_PAGE\_ SIZE respectively.  The idea behind using
  three different shifted versions of the same feature is that it
  allows us to look into a wider range of bits than with a single
  version.  It also helps give more importance to the overlapping bits
  and lesser importance to most and least significance bits.  This
  approach can also eliminate destructive interference that can be
  caused by directly folding the bits into half.

\item \textbf{Confidence}: The integer confidence on the scale of 0 to
  100 which was used in the original SPP design.  While the original
  confidence is not used directly for decision making, it can still
  contribute to the final decision made.

\item \textbf{Program Counter XOR Depth}: PC is for the instruction
  which triggered the prefetch chain.  Depth refers to the iteration
  count of the lookahead stages.  In general, PC is not considered as
  a good basis for doing lookahead prefetching as all the prefetches
  with depth >= 1 gets aliased onto the same PC which won't be true in
  actual demand fetch case.  This feature resolves a PC into a
  different value for each lookahead depth of SPP speculation, giving
  a more accurate correlation in lookahead cases.  This is akin to the
  concept of Virtual Program Counters\cite{VPC} introduced by Kim
  \textit{et. al.} for indirect branch prediction.

\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
    PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$
  PC before the instruction that triggered the current prefetch.
  Hashing together the last three PCs tell SPP-Perc about the path
  that led to the current demand access and helps capture and
  branching information of the current basic block.  PCs are shifted
  in the increasing order of history before being hashed together.
  This is done to avoid the resultant value of zero in case 2 or more
  PCs are the same.  Additionally, obfuscating the information as it
  gets old allows us to get a wider and an approximate look into the
  program history.

\item \textbf{Program Counter XOR Delta}: This tells us if a given PC
  favours particular value(s) of delta.  As noted earlier, while PC in
  itself does not much useful information, this hash resolves PC into
  different values based on the tendency of that PC to favour a
  certain delta.  This way, dynamic nature of different instances of
  the same memory access instruction can be captured here.

\item \textbf{Current Signature XOR Delta}: It can be recalled from
  the discussion of SPP in section XX that the new signature is
  generated using the old signature and the blobk delta.  While
  `Current Signature XOR Delta' is not the exact formula for
  generating the future signature, it gives an approximate idea of the
  path that the combination of these two values can lead to.

%%%%%%% Below are the features rejected when downsizing 14 -> 9
% \item \textbf{Current Signature}: The 12-bit signature used by SPP
%   to index into the Pattern Table
% \item \textbf{Delta}: The signed version of the difference in the
%   block address that triggers the prefetch and the block address of
%   the data being prefetched
% \item \textbf{PC}: Program Counter of the Load instruction that
%   triggers that particular prefetch
% \item \textbf{Base address XOR Delta}: If any particular value(s) of
%   delta are highly favoured by a given address which triggered the
%   prefetch, this composite feature can capture that information
% \item \textbf{Cache line XOR Delta}: qThis feature captures if any
%   particular delta is highly favoured by a given cache line address
%\item \textbf{Confidence XOR Delta}: This gives a combined insight
%  into that fact that certain delta-confidence pairqs predicted by
%  SPP might correlate more with the prefetch outcome
%%%%%%%%%%%%%%%%%%%%%%

\end{itemize}

Some of the composite features are derive from simple hashing (XOR) of
two primary features.  There is always a question of usefulness of
such composite features and the `fresh' information added.  We try to
justify the choice of each feature by quantifying the contribution
made towards predicting prefetch behaviour, in Section XX.  Finally,
as noted above, each feature indexes into its independent entry of
perceptron weights.