### 2018-12-01 intro.tex
reworded almost everything

\section{Introduction}
\label{Introduction}

Processor and memory technologies have been developed with different
goals in mind.  While processor scaling has focused on speed improvements,
memory scaling has primarily focused on increasing capacity.  The
difference in each technology's scaling has led to the Memory Wall~\cite{MemWall}
-- the increasing gap between processor and memory performance. Data prefetching
is one important technique that has been developed to minimize the effects 
of this trend.

%Data prefetching exploits the fact that in most applications,
% memory accesses have a repeating and predictable pattern to speculatively fetch useful data from slower levels
%of the cache hierarchy into faster higher levels of cache. %phrasing?

An ideal prefetching scheme would perfectly capture a program's memory access pattern, and then
predict and pre-load the needed data into the processor's caches in a timely manner. 
Memory access patterns may be simple, such as accessing every item in an array with a for-loop,
or very complex, such as chasing pointers through dynamically-allocated memory.
Practical prefetchers face challenges not only in predicting what data will be useful in the future,
but also in timing when data should be prefetched, deciding which level of the cache hierarchy
the data should be stored in, and deciding what data should be evicted from the caches to accomodate
the prefetched data.

All prefetchers are designed around a fundamental trade-off between two important metrics:
coverage and accuracy.  Prefetcher coverage refers to the number of cache hits that would have
been cache misses without the prefetcher.  For example, if an application experiences 1,000 cache misses
without a prefetcher, but only 200 cache misses with a prefetcher, then the prefetcher has 80\%
coverage for that application.  Prefetcher accuracy refers to the fraction of prefetched cache lines
that end up being used by the application.  So if a prefetcher prefetches 1,200 cache lines, 
but only 800 of them are used by the application, then that prefetcher's accuracy is 66.7\%.

Coverage and accuracy are generally at odds with one another, and as one metric improves, the other
usually gets worse.  For example, when an application accesses a new region of memory for the first
time, a na\"ive prefetcher may predict that all data in that region will be used by the application.
This will clearly result in 100\% coverage for that region, but with possibly a very low accuracy.
In fact, so much bandwidth may be wasted prefetching unused data that performance can ultimately be
harmed by this strategy.  On the other extreme, another prefetcher may be overly
conservative and never prefetch anything, wasting no bandwidth, and achieving 0\% prefetch
coverage.

A delicate balance between coverage and accuracy is required for a prefetcher to maximize
its performance impact.  Prefetchers are generally designed with internal mechanisms to
monitor their accuracy, with throttling thresholds that can be tuned for either coverage 
or accuracy.  The more irregular an application's memory access pattern is, the more difficult
it is to accurately predict every access, so a prefetcher will have to be tuned more toward
coverage (and away from accuracy) in order to gain any benefit.  This may be especially dangerous
to do in the context of a multi-core processor, where overly aggressive prefetching in one core can
waste shared resources, such as last-level cache capacity, and off-chip bandwidth~\cite{Friendly}.

%Predicting future accesses requires a trade-off between coverage, the percent of predictions that 
%prevented a cache miss, and accuracy, the percent of predictions made that were correct. 
%%Gino: Just added a short explanation of coverage and accuracy tradeoff, not sure if necessary
%A prefetcher may have high coverage by reducing the number of misses by simply requesting a large 
%amount of data, however this does not imply accuracy. Even if the prefetcher brought in data accessed in the future,
%a high percentage of the data requested could never be accessed meaning the prefetcher had low accuracy, 
%resulting in wasted resources as well as cache pollution. Likewise, a prefetcher can have a 
%higher accuracy by being completely sure the data requested is used in the future, 
%but may not affect performance since it did not have the coverage needed to make an impact. %end addition.
%Most prefetchers maintain an internal confidence mechanism by keeping counters
%to track events such as cache hits, misses, references to an entry in the prefetcher's
%structure, etc. By comparing the confidence indicated by the counters against different 
%threshold values, the aggressiveness of the prefetcher can be adjusted, resulting in lower
%coverage and higher accuracy.

%For a prefetcher to capture highly irregular access patterns, it needs to be
%highly aggressive with high coverage. This causes the prefetcher to recommend many
%low accurracy prefetches that lead to the cache being polluted with useless data.  
%This effect is more pronounced in a multi-core scenario where the last-level cache is a
%shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
%be highly accurate but might not make enough useful prefetches due to its lower coverage.

We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
existing state-of-the-art prefetchers.  PPF works by observing the stream of
candidate prefetches generated by a prefetcher, and then rejects those that
are predicted by the online-trained neural model to be inaccurate.
The state-of-the-art prefetcher that we 
use to evaluate PPF in this paper is the Signature Path Prefetcher 
(SPP)~\cite{SPP}, however in principle, PPF could benefit any prefetcher.
In this design, PPF replaces SPP's existing confidence-based throttling mechanism,
which itself was a highly tuned feature of that prefetcher.  Because PPF is
so much more effective at rejecting inaccurate prefetches than SPP's baseline
mechanism, we are free to re-tune the rest of SPP's design around maximizing
coverage.  The result is an increase in both accuracy and coverage, and
a notable increase in performance.

%We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
%the existing state of the art prefetchers to robustly overcome the coverage
%vs. accuracy trade-off.  The idea is to use a state-of-the-art prefetcher as
%the base engine. For that we choose Signature Path Prefetcher (SPP) and
%modify the design to make it as aggressive as possible.  This modification
%enables the base engine to capture complex memory access patterns and go
%deeper into the speculation path, increasing coverage without sacrificing accuracy. 
%Normally such aggressive prefetching would come at the cost of increased DRAM traffic
%and cache pollution. However, prefetch suggestions recommended by the base prefetch
%engine are passed through the perceptron based filter to ensure an accurate prediction.
%Over time, PPF learns to correlate the prefetch recommendations with various available
%features, leading to elimination of useless prefetches. This filtering leads
%to overall increased accuracy of the prefetch scheme and reduction in DRAM
%traffic and cache pollution.

\vspace{1ex}This paper describes PPF, explains its merits, offers analysis, and
outlines the scope for future research. Its contributions include:

\begin{itemize}

\item An on-line neural model used for hardware data
  prefetching.  Previous work in this area either relied on program
  semantics~\cite{Semantics} or were application specific~\cite{Datacenter}.
% djimenez: removing reference to this arXiv paper. we are not obligated to cite it.
%, or
%  could not yield a significant improvement over the baseline prefetcher
%  used~\cite{BadPerc}.

\item Implementing PPF in a robust and accurate
  practical prefetching mechanism, yielding a significant performance
  improvement compared to other state-of-the-art prefetchers.
  Moreover, PPF learns to adapt itself to shared resource
  constraints, leading to further increased performance in multi-core and
  bandwidth-constrained environments.

\item A methodology for determining an appropriate set of 
  features for prediction, regardless of the underlying prefetcher used.
  More details are explained in Section \ref{Method-Features}.

\end{itemize}

In a single core configuration, PPF increases performance by 6.84\%
compared to the baseline prefetcher, SPP.  In a multi-core system running a
mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement
of 11.9\% over SPP for a 4-core system, and 10.67\% for an 8-core system.

The paper is organized as follows. Section~\ref{Background} describes
background work and summarizes SPP.  Section~\ref{Arch} gives the 
architectural overview of PPF. Section~\ref{Impl} discusses the
implementation of PPF using SPP and the particular features used.
Section~\ref{Method} gives the evaluation methodology and explores the feature
space for perceptron learning.  Section~\ref{Results} evaluates the
performance of the prefetcher and Section~\ref{Conclusion} concludes the
paper.