### 2018-12-04 design_impl.tex
Changes in Design. Minor edits in intro / method / results

\section{PPF Implementation Using SPP}
\label{Impl}

This section describes our implementation of PPF and the range of
features that are used by the perceptron for correlating the
prefetches.  For a practical implementation, we have used SPP as the
base engine prefetcher.

\subsection{Changes made to original SPP}
\label{Impl-Changes}
To modify the SPP design to suit our scheme, the following changes
were made:

\begin{itemize}
\item \textbf{Original Thresholds discarded}\newline In PPF, the
  perceptron sum is used to decide whether to prefetch or not, and the
  fill-level in case of prefetch.  Thus, the confidence thresholds --
  T\textsubscript{f} and T\textsubscript{p} are no longer needed.

% EB: Can this point be removed?
% djimenez: I think this would be necessary to reproduce our results so leave it in. 
% EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.

%\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
%  original SPP does not consider the slots available in the L2 MSHR
%  queue.  It has the assumption that at no time can the number of
%  suggested prefetches exceed the capacity of the queue.  While it was
%  not described in the paper, the internal confidence mechanism makes
%  sure that the above assumption was maintained.  In PPF, since there
%  cannot be an assumption on the degree of aggressiveness of the
%  underlying prefetcher, we explicitly check that at no point should
%  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
%    this a big deal?  Seems like a detail}

\item \textbf{Enhanced Prefetch Table}\newline The prefetch table
  introduced in the original SPP keeps track of up to 1024 L2 cache
  prefetch suggestions.  For PPF, the prefetch table is modified to
  store the features used for perceptron indexing..
  This metadata is needed to retrieve the
  values of the features to train the perceptron at a later stage in
  time. The prefetch table mentioned above keeps track of the
  prefetches that actually happened in the L2 Cache.  We introduce the
  reject table to keep a track of the prefetches that were suggested
  by the base SPP engine but were rejected by PPF.  In case it is
  detected that a future demand fetch could have used this rejected
  prefetch, the perceptron weights are updated to reflect that
  misprediction.
\end{itemize}

\subsection{Features used by Perceptron}
\label{Impl-Features}
Here we discuss the various features that correlate the prefetching
decision with the program behavior. All the features we used can be 
derived from the information available in the L2 Cache access stream.
Our feature selction involved searching over a large space of 
relevant perceptron features. Using the statistical methodology 
outlined in Section~\ref{Method-Features}, we pruned the feature 
set to a minimal yet relevant set of features
% \pg{again you want to get at why
% these were chosen.  Also mention that you searched over a large
% space of possible features and these were the best.}

% [EB] Tried addressing. Does it look reduntant when read after 'Feature Selection'
% bullet from the previous section?

\begin{itemize}
\item \textbf{Base Address}: The base address of the demand 
access that triggers the prefetch coresponds to a stream of 
future accesses that SPP and the PPF have seen before. 
Therefore, PPF will correlate the past behavior of this address 
to the confidence of a prefetch.

\item \textbf{Cache Line} and \textbf{Page Address}: These two
  separate features are derived from the base address that triggered
  the prefetch, as follows: base\_addr >> LOG2\_BLOCK \_SIZE and
  base\_addr >> LOG2\_PAGE\_SIZE respectively. The idea behind using
  three different shifted versions of the same feature is that it
  allows us to look into a wider range of bits than with a single
  version.  It also helps give more importance to the overlapping bits
  and lesser importance to most and least significance bits. This
  approach can also eliminate destructive interference that can be
  caused by directly folding the address bits into half.

% djimenez: this is a nice intuition and explanation. 
% [EB] Thank you!

\item \textbf{Program Counter XOR Depth}: The PC is for the
  instruction that triggered the prefetch chain.  Depth refers to the
  iteration count of the lookahead stages.  In general, the PC is not
  considered as a good basis for doing lookahead prefetching as all
  the prefetches with depth >= 1 are aliased into the same PC which
  is not true in an actual demand access.  This feature resolves
  a PC into a different value for each lookahead depth of prefetch
  speculation, giving a more accurate correlation in lookahead cases.
  This is akin to the concept of Virtual Program Counters~\cite{VPC}
  introduced by Kim \textit{et. al.} for indirect branch prediction.

\item \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
    PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$
  PC before the instruction that triggered the current prefetch.
  Hashing together the last three PCs tell PPF about the path that led
  to the current demand access and helps capture and branching
  information of the current basic block.  PCs are shifted in the
  increasing order of history before being hashed together.  This is
  done to avoid the resultant value of zero when 2 or more PCs are the
  same.  Additionally, obfuscating the information as it gets old
  allows us to get a wider and approximate look into the program's
  history.

\item \textbf{Program Counter XOR Delta}: This feature tells us if a
  given PC favours particular value(s) of delta.  As noted earlier,
  while the PC alone does not convey useful information, this
  hash resolves the PC into different values based on the tendency of
  that PC to favour a certain delta.  Thus, the dynamic nature of
  different instances of the same memory access instruction is 
  captured here.
 % GAC: Seems more predictable than dynamic if its capturing 
 % what is essentially a streaming access pattern. Not quite sure how to rephrase,
 % if I'm correct

 % GAC: Again having trouble phrasing the below point. I was attempting to say
 % that despite SPP having low confidence due to speculation, PPF may 
 % reinforce prefetches with lower confidence if SPP has had a tendancy to be 
 % correct in the past.
\item \textbf{Confidence}: The integer confidence, on a scale of 0 to
  100, that was used in the original SPP design. While the original 
  confidence does not directly make the decision to prefetch,
  PPF correlates it to the correctness of a proposed prefetch.
  While the original SPP may have dismissed a prefetch due to running
  further into speculation, PPF can use the original confidence as indicator
  of not only when prefetches become less confident, but also how likely
  a low confident speculation is correct in the context of other features.

\item \textbf{Page Address XOR Confidence}: This feature scores the
  tendency of each page to be prefetch friendly or prefetch averse. It
  helps resolve a page into different entries depending on its
  confidence for prefetching, which can vary during phases of a
  program execution.

\item \textbf{Current Signature XOR Delta}: Recall from the discussion
  of SPP in Section \ref{Background-SPP} that the new signature is
  generated using the old signature and the current block delta.
%Gino: Added the sentence below, is this correct?
%[EB]: It gives the next signature instead of next delta
  The result of this feature is the next signature that is predicted to be
  accessed based on the delta predicted by SPP. While ``Current
  Signature XOR Delta'' is not the exact formula for generating the
  future signature, it gives an approximate idea of the path that the
  combination of these two values can lead to.

%%%%%%% Below are the features rejected when downsizing 14 -> 9
% \item \textbf{Current Signature}: The 12-bit signature used by SPP
%   to index into the Pattern Table
% \item \textbf{Delta}: The signed version of the difference in the
%   block address that triggers the prefetch and the block address of
%   the data being prefetched
% \item \textbf{PC}: Program Counter of the Load instruction that
%   triggers that particular prefetch
% \item \textbf{Base address XOR Delta}: If any particular value(s) of
%   delta are highly favoured by a given address which triggered the
%   prefetch, this composite feature can capture that information
% \item \textbf{Cache line XOR Delta}: qThis feature captures if any
%   particular delta is highly favoured by a given cache line address
%\item \textbf{Confidence XOR Delta}: This gives a combined insight
%  into that fact that certain delta-confidence pairqs predicted by
%  SPP might correlate more with the prefetch outcome
%%%%%%%%%%%%%%%%%%%%%%

\end{itemize}

Some of the composite features are derived from simple hashing (XOR)
of two primary features.  There is always a question of usefulness of
such composite features and the new information added.  We justify the
choice of each feature by quantifying the contribution made towards
predicting prefetch behaviour, in Section \ref{Method-Features}.
Finally, as noted above, each feature indexes into its independent
entry of perceptron weights.

% [EB]: Think of more implementation specific details
% 7 to 12 bits of indexing