### 2018-12-05 intro.tex

\section{Introduction}
\label{Introduction}

Processor and memory technologies have been developed with different
goals in mind.  While processor scaling has focused on speed improvements,
memory scaling has primarily focused on increasing capacity.  The
difference in each technology's scaling has led to the Memory Wall~\cite{MemWall}
-- the increasing gap between processor and memory performance. Data prefetching
is one important technique that has been developed to minimize the effects 
of this trend.

%Data prefetching exploits the fact that in most applications,
% memory accesses have a repeating and predictable pattern to speculatively fetch useful data from slower levels
%of the cache hierarchy into faster higher levels of cache. %phrasing?

An ideal prefetching scheme would perfectly capture a program's memory access pattern, and then
predict and pre-load the needed data into the processor's caches in a timely manner. 
Memory access patterns may be simple, such as accessing every item in an array with a for-loop,
or very complex, such as chasing pointers through dynamically-allocated memory.
Practical prefetchers face challenges not only in predicting what data will be useful in the future,
but also in timing when data should be prefetched, deciding which level of the cache hierarchy
the data should be stored in, and deciding what data should be evicted from the caches to accommodate
the prefetched data.

All prefetchers are designed around a fundamental trade-off between two important metrics:
coverage and accuracy.  Prefetcher coverage refers to the number of cache hits that would have
been cache misses without the prefetcher.  For example, if an application experiences 1,000 cache misses
without a prefetcher, but only 200 cache misses with a prefetcher, then the prefetcher has 80\%
coverage for that application.  Prefetcher accuracy refers to the fraction of prefetched cache lines
that end up being used by the application.  So if a prefetcher prefetches 1,200 cache lines, 
but only 800 of them are used by the application, then that prefetcher's accuracy is 66.7\%.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{Motivation}
\caption{Variation of IPC, Total Prefetches and Useful Prefetches}
\label{Fig:Motivation}
\end{figure}

Coverage and accuracy are generally at odds with one another, and as one metric improves, the other
usually gets worse.  For example, when an application accesses a new region of memory for the first
time, a na\"ive prefetcher may predict that all data in that region will be used by the application.
This will clearly result in 100\% coverage for that region, but with possibly a very low accuracy.
In fact, so much bandwidth may be wasted prefetching unused data that performance can ultimately be
harmed by this strategy.  On the other extreme, another prefetcher may be overly
conservative and never prefetch anything, wasting no bandwidth, and achieving 0\% prefetch
coverage.

Figure~\ref{Fig:Motivation} illustrates the above scenario in a working manner.
Here we consider a state-of-the-art lookahead prefetcher -- SPP. We keep on 
re-tuning its thresholds, each time getting a variant with an increasing lookahead 
depth coverage. The figure depicts the behavior of the prefetcher variants running 
on the trace {\tt 603.bwaves\_s-1740B}. The IPC, total number of prefetches 
suggested by the prefetcher (TOTAL\_PF) and the actual useful predictions (GOOD\_PF),
all have been normalized to the least aggressive version of the prefetcher. 
As the look-ahead depth of the prefetch suggestions increase, so do useful prefetches. 
But that comes at a cost of total prefetches increasing at an even higher rate. This leads 
to cache pollution and bandwidth contention, as is clearly visible with the reduction in 
the IPC. Similar behavior was observed in 12 other SPEC CPU 2017 simpoints.

Therefore, a delicate balance between coverage and accuracy is required for a prefetcher to maximize
its performance impact.  Prefetchers are generally designed with internal mechanisms to
monitor their accuracy, with throttling thresholds that can be tuned for either coverage 
or accuracy.  The more irregular an application's memory access pattern is, the more difficult
it is to accurately predict every access, so a prefetcher will have to be tuned more toward
coverage (and away from accuracy) in order to gain any benefit.  This may be especially dangerous
to do in the context of a multi-core processor, where overly aggressive prefetching in one core can
waste shared resources, such as last-level cache (LLC) capacity, and off-chip bandwidth~\cite{Friendly}.

%Predicting future accesses requires a trade-off between coverage, the percent of predictions that 
%prevented a cache miss, and accuracy, the percent of predictions made that were correct. 
%%Gino: Just added a short explanation of coverage and accuracy tradeoff, not sure if necessary
%A prefetcher may have high coverage by reducing the number of misses by simply requesting a large 
%amount of data, however this does not imply accuracy. Even if the prefetcher brought in data accessed in the future,
%a high percentage of the data requested could never be accessed meaning the prefetcher had low accuracy, 
%resulting in wasted resources as well as cache pollution. Likewise, a prefetcher can have a 
%higher accuracy by being completely sure the data requested is used in the future, 
%but may not affect performance since it did not have the coverage needed to make an impact. %end addition.
%Most prefetchers maintain an internal confidence mechanism by keeping counters
%to track events such as cache hits, misses, references to an entry in the prefetcher's
%structure, etc. By comparing the confidence indicated by the counters against different 
%threshold values, the aggressiveness of the prefetcher can be adjusted, resulting in lower
%coverage and higher accuracy.

%For a prefetcher to capture highly irregular access patterns, it needs to be
%highly aggressive with high coverage. This causes the prefetcher to recommend many
%low accuracy prefetches that lead to the cache being polluted with useless data.  
%This effect is more pronounced in a multi-core scenario where the LLC is a
%shared resource~\cite{Friendly}.  Conversely, a conservative prefetcher would
%be highly accurate but might not make enough useful prefetches due to its lower coverage.

We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
existing state-of-the-art prefetchers.  PPF works by observing the stream of
candidate prefetches generated by a prefetcher, and then rejects those that
are predicted by the online-trained neural model to be inaccurate.
The state-of-the-art prefetcher that we 
use to evaluate PPF in this paper is the Signature Path Prefetcher 
(SPP)~\cite{SPP}, however in principle, PPF could benefit any prefetcher.
In this design, PPF replaces SPP's existing confidence-based throttling mechanism,
which itself was a highly tuned feature of that prefetcher.  Because PPF is
so much more effective at rejecting inaccurate prefetches than SPP's baseline
mechanism, we are free to re-tune the rest of SPP's design around maximizing
coverage.  The result is an increase in both accuracy and coverage, and
a notable increase in performance.

%We propose Perceptron-based Prefetch Filtering (PPF) as an enhancement to
%the existing state of the art prefetchers to robustly overcome the coverage
%vs. accuracy trade-off.  The idea is to use a state-of-the-art prefetcher as
%the base engine. For that we choose Signature Path Prefetcher (SPP) and
%modify the design to make it as aggressive as possible.  This modification
%enables the base engine to capture complex memory access patterns and go
%deeper into the speculation path, increasing coverage without sacrificing accuracy. 
%Normally such aggressive prefetching would come at the cost of increased DRAM traffic
%and cache pollution. However, prefetch suggestions recommended by the base prefetch
%engine are passed through the perceptron based filter to ensure an accurate prediction.
%Over time, PPF learns to correlate the prefetch recommendations with various available
%features, leading to elimination of useless prefetches. This filtering leads
%to overall increased accuracy of the prefetch scheme and reduction in DRAM
%traffic and cache pollution.

This paper describes PPF, explains its merits, offers analysis, and
outlines the scope for future research. Its contributions are:

\begin{itemize}

\item An on-line neural model used for hardware data prefetching.  Previous
work in this area either relied on program semantics~\cite{Semantics} or were
application specific~\cite{Datacenter}.

\item Implementing PPF filtering a state-of-the-art prefetchier, giving a
significant performance improvement compared to previous work.  PPF learns to
adapt itself to shared resource constraints, leading to further increased
performance in multi-core and bandwidth-constrained environments.

\item A methodology for determining an appropriate set of features for
prediction, regardless of the underlying prefetcher used.  More details are
explained in Section \ref{Method-Features}.

\end{itemize}

In a single core configuration, PPF increases performance by 6.84\%
compared to the underlying prefetcher, SPP.  In a multi-core system running a
mixes of memory intensive SPEC CPU 2017 traces, PPF saw an improvement
of 11.9\% over SPP for a 4-core system, and 10.67\% for an 8-core system.

% djimenez - cutting this
%The paper is organized as follows. Section~\ref{Background} describes
%background work and summarizes SPP.  Section~\ref{Arch} gives the 
%architectural overview of PPF. Section~\ref{Impl} discusses the
%implementation of PPF using SPP and the particular features used.
%Section~\ref{Method} gives the evaluation methodology and explores the feature
%space for perceptron learning.  Section~\ref{Results} evaluates the
%performance of the prefetcher and Section~\ref{Conclusion} concludes the
%paper.