### 2018-12-07 conc.tex

\section{Conclusion}
\label{Conclusion}
In this paper, we introduce perceptron-based learning for prefetching. The
perceptron filter acts independently as a check on the underlying prefetch
engine. PPF improves performance over the base prefetcher by up to 11.4\%. 
We believe that this is a robust and adaptable technique that can be used 
to enhance any existing and upcoming prefetchers.