### 2018-12-07 method.tex
finished pass

\section{Methodology}
\label{Method}

\subsection{Performance Model}
\label{Method-Model}
We use the ChampSim~\cite{Champsim} simulator for the evaluation of
PPF against prior work techniques.  ChampSim is an enhanced version of
the framework that was used for the 2nd Data Prefetch Championship
(DPC-2)~\cite{DPC_2}, also used in the 2nd Cache Replacement
Competition (CRC2)~\cite{CRC_2}.  We model 1-core, 4-core, and 8-core
out-of-order machines.  The details of the configuration parameters
are summarized in Table~\ref{tab:Sim_params}.

\begin{table}[]
    \centering
    \begin{adjustbox}{width=0.9\columnwidth}
    \begin{tabular}{|c|c|}
    \hline
    	 Block & Configuration\\
    \hline
	 \multirow{2}{1.5cm}{CPU Core} 		& 1-8 Cores, 4 GHz\\
						& 256 entry ROB, 4-wide\\
    \hline
         \multirow{2}{2.7cm}{Private L1 DCache} & 32 KB, 8-way, 4 cycles\\
						& 8 MSHRs, LRU\\
    \hline
         \multirow{2}{2.4cm}{Private L2 Cache}  & 256 KB, 8-way, 8 cycles\\
						& 16 MSHRs, LRU, Non-inclusive\\		
    \hline
         \multirow{2}{1.7cm}{Shared LLC} 	& 2MB/core, 16-way, 12 cycles\\
						& 32 MSHRs, LRU, Non-inclusive\\
    \hline
         \multirow{3}{1.1cm}{DRAM} 		& 4 GB 1-Channel (single-core)\\ 
						& 8 GB 2-Channels (multi-core)\\ 
						& 64-bit channel, 1600MT/s\\
    \hline
    \end{tabular}
    \end{adjustbox}
    \caption{Simulation Parameters}
    \label{tab:Sim_params}
\end{table}

% djimenez: changing all the "was" to "is." when describing our work, use
% present tense

The block size is fixed at 64 bytes.  Prefetching is only triggered
upon L2 cache demand accesses but could be directed to the L2 or
last-level cache.  No L1 data level prefetching is done. The LRU
replacement policy is used on all levels of cache hierarchies. Branch
prediction is done using the perceptron branch
predictor~\cite{PerceptronPredictor}.  The page size is 4KB.  ChampSim
operates all the prefetchers strictly in the physical address space.

\subsection{Testing Under Additional Memory Constraints}
\label{Method-AdditionalMem}
The default single-core configuration simulates a 2MB LLC and a single
channel DRAM with 12.8GB/s bandwidth. We extend the simulations to
include memory constraints introduced in DPC-2. Specifically we look
at the following two variations: Low Bandwidth DRAM, where DRAM
bandwidth is limited to 3.2 GB/s, and small LLC, where the LLC size is
reduced to 512 KB.  All the multi-core simulations are only done in
the default configuration.

\subsection{Workloads}
\label{Method-Workloads}
% djimenez: don't say this is the first time SPEC CPU 2017 has been used this way.
% i don't know if that's true, and in any event the value is small.

%This is the first time that SPEC CPU 2017 benchmark suite~\cite{SPEC2017}
%has been used to characterize and measure the prefetch performance.
%We use the SPEC CPU 2017 benchmark suite~\cite{SPEC2017}. 
We use all the 20 workloads available in the SPEC CPU 2017
suite~\cite{SPEC2017}. Using the SimPoint~\cite{SimPoint} methodology,
we identified 95 different program segments of 1 Billion
instructions each.

\noindent \textbf{Single-core performance:} For single-core
simulations, we use the first 200 million instructions to warm-up the
microarchitectural structures and the next one billion instructions to
do detailed simulations and collect run-time statistics. We report the
IPC speedup over the baseline of no prefetching.  The final numbers
reported are the geometric mean of the weighted mean speedup achieved
per application using the SimPoint methodology.

\noindent \textbf{Multi-core performance:} For multi-application
workloads, we generate 100 random mixes and another 100 mixes from the
memory intensive subset of SPEC CPU 2017.  For 4-core workloads, 200
Million instructions are used for warm-up and additional 1 Billion
instruction simulated for collecting statistics.  Each CPU keeps
executing its workload till the last CPU completes one billion
instructions after warm-up.  For collecting IPC and other data, only
the first billion instructions are considered as the region of
interest.

Here we report the weighted speedup normalized to baseline
\textit{i.e.}, no prefetching. For each of the workloads running on a
particular core of the 4-core 8 MB LLC system, we compute
IPC\textsubscript{i}. We then find the IPC\_isolated\textsubscript{i}
of the same workload running in isolated 1-core 8 MB LLC environment.
Then we calculate the total weighted-IPC for a given workload mix as
$\Sigma$ (IPC\textsubscript{i} / IPC\_isolated\textsubscript{i}). For
each of the 100 workload-mix, the sum obtained is normalized to the
weighted-IPC calculated similarly for baseline case \textit{i.e.}, no
prefetching, to get the weighted-IPC-speedup. Finally the geometric
mean of these 100 weighted-IPC-speedup is reported as the effective
speedup obtained by the prefetching scheme.

We repeat the same process for 8-core workloads, correspondingly with
16MB LLC. The only difference is that 20 million warm-up instructions
and 100 million full instructions are executed. This is done so as to
keep the simulation run-time within reasonable limits as a single
8-core mix takes up to 3 days to simulate one billion instructions.

\noindent \textbf{Validation:} We cross-validated our PPF model using
SPEC CPU 2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite}
benchmarks. For single-core SPEC CPU 2006, we developed 94 simpoints
spread across all the 29 applications.  For multi-core, we followed
the same methodology as SPEC CPU 2017.  For CloudSuite, we used the
traces made available for the 2nd Cache Replacement Competition
(CRC-2)~\cite{CRC_2}. The traces include four 4-core applications with
six distinct phases per application.

In total, we used 285 traces representing workloads across 53
applications.  Throughout the paper, we consider memory intensive
subset as the applications with SimPoint weighted LLC MPKI > 1. This
includes 11 out of 20 SPEC CPU 2017 applications. For SPEC CPU 2006,
this includes 16 out of 29 applications.

\subsection{Prefetchers Simulated}
\label{Method-Prefetchers}
We compared PPF against three of the latest, state of the art
hardware-only prefetchers: Best Offset Prefetcher (BOP), DRAM Aware -
Access Map Pattern Matching (DA-AMPM) \cite{DA_AMPM} and Signature
Path Prefetcher (SPP).  BOP was the winner of 2nd Data Prefetching
Championship. DA-AMPM is the enhanced version of AMPM, modified to
account for DRAM row buffer locality. SPP has been shown to outperform
BOP on SPEC CPU 2006 traces~\cite{SPP}. For each of these, we compare
their speedups taking the no prefetching case as the baseline.

\subsection{Developing Features for PPF}
\label{Method-Features}
This section describes the intuition and analysis that went behind
developing the perceptron features.  As noted earlier, we developed a
set of nine features that allow the perceptron layer to correlate
prefetching decision with the program behavior.  To study the
correlation across each feature, we statistically examine the
perceptron weights and try to interpret their distribution.

\noindent \textbf{Global Pearson's Correlation} Here we examine the
perceptron weights at the end of all trace execution by which time the
weights have settled to steady values. The weights obtained from
running all the SPEC CPU 2017 traces are concatenated.  Features with
a bulk of their perceptron weights concentrated around 0 or small
magnitude numbers show a weak correlation with the prefetching
outcome.  On the other hand, features with most of the weights
saturated around highest value (+15) show a high positive correlation
and the features with weights close to the lowest value (-16) show a
strong negative correlation.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\columnwidth]{Hist}
    \caption{Distribution of Trained Weights}
    \label{Fig:Hist}
  \end{center}
\end{figure}

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\columnwidth]{Pvals}
    \caption{P-Values for all Features}
    \label{Fig:Pvals}
  \end{center}
\end{figure}

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\columnwidth]{Pvals_Trace}
    \caption{P-value Variation across Traces for selected Features}
    \label{Fig:Pvals_Trace}
  \end{center}
\end{figure}

We plot a histogram for each feature depicting weights distribution
from -16 to +15 and generate the Pearson's correlation factor for that
feature.  Pearson's factor is a numerical measure ranging from -1 to 1
of the degree of linear correlation between two variables. The
magnitude of Pearson's factor gives the extent of correlation and the
sign indicates whether it is a positive correlation or a negative
correlation. Values close to 0 suggest a low correlation while a value
of +1/-1 suggests a perfectly linear positive / negative correlation
respectively.

As a part of our perceptron feature selection methodology, we explored
a wide variety of features to begin with. Features with a low
Pearson's coefficient were rejected as they didn't provide much useful
correlation.  Figure~\ref{Fig:Hist} depicts the histogram distribution
of trained weights for two features.  The first feature,
\textit{Confidence XOR Page address} has the highest observed P-value,
hence was retained.  On the other hand, the second feature,
\textit{Last Signature} did not provide any meaningful correlation and
hence was rejected.  This is visible from the bulk of its trained
weights settling to near zero values.

Figure~\ref{Fig:Pvals} shows all the features which are finally used,
arranged in the increasing order of their Pearson's factor.  As can be
seen 5 out of the 9 features provide a moderate to high correlation,
with the magnitude of P-value > 0.6. The single most important
feature, \textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90.

\noindent \textbf{Per Trace Correlation} Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces. Here we give special attention to features
with low P-values in. Figure~\ref{Fig:Pvals_Trace} shows the variation
of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and \textit{PC XOR Depth}; across all the
SPEC CPU 2017 traces.  For simplicity, the traces are arranged in an
increasing order of contribution made by the feature.  It can be seen
that even features with a low overall correlation provide useful
correlation (magnitude > 0.5) for a significant number of traces. This
study motivated us to choose \textit{PC XOR DELTA} over \textit{Last
  Signature} as it provided useful correlation in at least some of the
traces.

\noindent \textbf{Trimming Features Using Cross-correlation} Beside
providing interesting insights into prefetching behavior, P-value can
also be used for feature selection and prefetcher tuning.

As we examined correlation of each feature with the final outcome, we
also studied correlation between the features.  We used the above
methodology to eliminate features providing little information that
has already been captured in other features.

We initially came up a set of 23 features.  By studying cross
correlation of each of these features against others in a 23$\times$23
matrix, we identified pairs of features with correlation factor > 0.9
in magnitude and eliminated redundant features, using guidance from
Global and per-trace Pearson's factor of those features.  By doing
this, we reduced the feature count to 9.  Thus, in the final
implementation of PPF, no two features have a high correlation between
them. This way we can be sure that each feature makes a contribution
that cannot be captured using other features.

Secondly, studying the relative importance of each feature enabled us
to vary the number of entries dedicated for each feature. Features
with higher correlation, \textit{cache line} and \textit{page address}
were given most importance and allowed full 12-bits of indexing.
Features like \textit{PC XOR delta} and \textit{PC XOR depth} with a
low overall P-value were allocated fewer entries in the feature table.

%To conclude, in the above discussion we justify the features for
%perceptron from a statistical viewpoint.  We also show how this
%information can also be used for prefetcher tuning.  All this study
%was made possible only because we used on-line perceptron learning for
%prefetching and that enabled us to examine the weights in detail.

\begin{table}[ht]
    \centering
    \begin{adjustbox}{width=0.9\columnwidth}
    \begin{tabular}{|c|c|c|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid 		& 1  & Indicates a valid entry in the table\\
         Tag 		& 6  & Identifier for the entry in the table\\
         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
                    	&    & useful demand fetch\\
         Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
    \hline
        PC 		& 12 & \\
        Address 	& 24 & \\
        Curr Signature 	& 10 & \multirow{2}{4.8cm}{Meta-data required for perceptron}\\
	PC$_i$ Hash	& 12 & \multirow{2}{1.1cm}{training}\\
        Delta 		& 7  & \\
        Confidence 	& 7  & \\
	Depth		& 4  & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 85 bits}}\\
    \hline
    \end{tabular}
    \end{adjustbox}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}

\subsection{Overhead for PPF}
\label{Method-Overheads}
In this section, we analyze the hardware overhead required to
implement PPF. The Prefetch Table was enhanced to accommodate storing
of meta-data for perceptron training. Table \ref{tab:PTable_metadata}
depicts the meta-data stored for each entry in the Prefetch Table.
Table \ref{tab:PPF_overhead} shows the total storage overhead of PPF
implementation. The hardware budget for 2nd Data Prefetching
championship was 32 KB.  Keeping that in mind the considerable speedup
PPF obtained over the winner, the extra hardware budget can be
accounted for. The extra hardware also makes the overall scheme more
scalable than SPP. In the original SPP paper, it was demonstrated that
adding extra hardware brings little advantage in terms of performance
gain. The newly added perceptron tables can be scaled to increase /
decrease features depending on the permitted budget.

\begin{table}[h]
\begin{adjustwidth}{}{}
    \centering
    \begin{adjustbox}{width=0.8\columnwidth}
    \begin{tabular}{|c|c|c|c|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
                                            &  \multirow{5}{0.5cm}{256}    & Valid (1 bit)  &             \\
                                             &      & Tag (16 bits)        &  \multirow{2}{0.9cm}{11008}           \\
                            Signature Table  &   & Last Offset (6 bits) &  \multirow{2}{0.5cm}{bits}  \\  
                                             &      & Signature (12 bits)  &             \\
                                             &      & LRU (6 bits)         &             \\
    \hline
                                    &  \multirow{3}{0.5cm}{512}    & $C_{sig}$ (4bits)      &\multirow{2}{0.9cm}{24576}               \\
                       Pattern Table         &   & $C_{delta}$ (4*4 bits) &  \multirow{2}{0.5cm}{bits}  \\
                                             &      & Delta (4*7 bits)       &               \\
    \hline
        \multirow{4}{1.5cm}{Perceptron\newline}     & 4096*4    & \multirow{4}{0.8cm}{5 bits}  & \multirow{3}{1.1cm}{113280}             \\
        \multirow{3}{1.2cm}{Weights}                & 2048*2    &           &  \multirow{3}{0.5cm}{bits}  \\
                                                    & 1024*2    &           &               \\
                                                    & 128*1     &           &              \\
    \hline
        Prefetch                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{85 bits}       & 87040 \\
        Table\footnotemark[1]   &           &               & bits\\
    \hline
        Reject                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{84 bits}    & 86016 \\
        Table\footnotemark[2] & & & bits\\
    \hline
        \multirow{4}{1.0cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
        \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
        \multirow{2}{1.2cm}{Register}               &                        & Last Offset (6 bits) &                               \\
                                                    &                        & Delta (7 bits)       &                               \\
    \hline
        Accuracy        & 1     & C$_{total}$       & 10 bits   \\
        Counters        & 1     & C$_{useful}$      & 10 bits   \\
    \hline
        \multirow{3}{1.5cm}{Global PC\newline}      &       & $PC_1$ (12 bits)      &           \\
        \multirow{2}{1.5cm}{~Trackers}              & 3     & $PC_2$ (12 bits)      & 36 bits   \\
                                                    &       & $PC_3$ (12 bits)      &           \\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 3,222,940 bits = 39.34 KB}}\\
    \hline
    \end{tabular}
    \end{adjustbox}
    \caption{SPP-Perc Storage Overhead}
    \label{tab:PPF_overhead}
\end{adjustwidth}
\end{table}

\footnotetext[1]{Components of Prefetch Table can be found in Table
  \ref{tab:PTable_metadata}.}

\footnotetext[2]{The Reject Table does not need to maintain the useful
  bit as that only applies for prefetches that ultimately made
  through.}

In terms of computations, the perceptron mechanism only introduces an
extra adder tree. The hash perceptron mechanism makes sure than there
is no actual vector multiplication happening in the hardware.
Obtaining the perceptron sum requires addition of 9 5-bit numbers.
Using an adder tree of 4 5-bit adders, this can be done in
ceil($log_{2}9$) = 4 steps. Perceptron update only requires weight
update by +1 or -1. Thus, all the operations required for perceptron
inferencing or updating the states of the perceptrons can be easily
done in the time constraints of L2 Cache Accesses.