### 2019-03-31 conc.tex

\section{Conclusion}
\label{Conclusion}
In this paper, we introduce the Perceptron-Based Prefetch Filtering
(PPF).  PPF acts as an independent check on the quality of predictions
made by the underlying prefetch engine.  We also created a case study
implementation of PPF using SPP as the {\color{red}underlying prefetcher}, while
in principle other prefetchers could be used.  We show that PPF
effectively filters bad prefetches, such that the given underlying
prefetcher can be highly aggressively tuned to achieve increasing
coverage.  PPF improves performance over the {\color{red}underlying} 
prefetcher by up to 11.4\%.  PPF is a robust and adaptable technique 
that can be used to enhance any existing prefetcher and can be a 
valuable tool in the design of future memory latency constrained systems.