### 2019-04-17 design_impl.tex

\section{PPF Implementation Using SPP}
\label{Impl}

This section describes a case study implementation of PPF and the
range of features that are used to determine the usefulness of
prefetches.  Here, we have selected SPP as the underlying prefetcher.

\subsection{Changes Made to SPP}
\label{Impl-Changes}
\noindent To modify the SPP design to suit our scheme, the following changes
were made:
\newline
\newline
\noindent \textbf{Exporting Features from SPP:} PPF uses the metadata 
specific to SPP, to build some of the perceptron features. These include
the lookahead depth, signature and the confidence counter. These features 
were made visible to PPF.
%
\newline
\newline
\noindent \textbf{Original Thresholds Discarded:} In PPF, the
perceptron sum is used to decide whether to prefetch or not, and the
fill-level in case of prefetch.  Thus, the confidence thresholds used 
by SPP -- T\textsubscript{f} and T\textsubscript{p} are no longer needed to
throttle the prefetcher and can be discarded.
%
%\pg{Note, however, as we discuss later, we
%use these confidence thresholds a feature in PPF so the mechanism is
%retained. -- right?}
% djimenez: keeping comments on every line here so it doesn't screw up my
%"roll-your-own" itemized list. don't put in any newlines.
%
%
% EB: Can this point be removed?
% djimenez: I think this would be necessary to reproduce our results so leave it in. 
% EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.
%
%\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
%  original SPP does not consider the slots available in the L2 MSHR
%  queue.  It has the assumption that at no time can the number of
%  suggested prefetches exceed the capacity of the queue.  While it was
%  not described in the paper, the internal confidence mechanism makes
%  sure that the above assumption was maintained.  In PPF, since there
%  cannot be an assumption on the degree of aggressiveness of the
%  underlying prefetcher, we explicitly check that at no point should
%  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
%    this a big deal?  Seems like a detail}
%

%% \pg{since you actually moved this table into PPF, most this discussion
%%   should actually go in the PPF arch section, I think its generic to
%%   all PPFs likely anyways.  Instead here I'd talk about how you
%%   removed the prefetch filter that was part of SPP since its redundant
%%   with the function of PPF}
%% \noindent \textbf{Enhanced Prefetch Table:} The prefetch table
%% introduced in the original SPP keeps track of up to 1024 L2 cache
%% prefetch suggestions. For PPF, the prefetch table is modified to store
%% the features used for perceptron indexing.  This metadata is needed to
%% retrieve the values of the features to train the perceptron at a later
%% stage in time. The prefetch table mentioned above keeps track of the
%% prefetches that actually happened in the L2 Cache.  We introduce the
%% reject table to keep a track of the prefetches that were suggested by
%% the base SPP engine but were rejected by PPF. In case it is detected
%% that a future demand fetch could have used this rejected prefetch, the
%% perceptron weights are updated to reflect that misprediction.

\subsection{Features used by Perceptron}
\label{Impl-Features}
Here we discuss the various features that correlate the prefetching
decision with the program behavior.  All the features we used can be
derived from the information available in the L2 Cache access stream
or are taken as metadata derived from the {underlying} prefetcher.  Our
feature selection involved searching over a large space of relevant
perceptron features.  Note that part of the process of tuning PPF to a
specific prefetcher involves examining the available metadata in the
prefetcher itself, and thus PPF is attuned to the {underlying}
prefetcher's design. Using the statistical methodology outlined in
Section~\ref{Method-Features}, we pruned the feature set to a minimal
yet relevant set of features.
% \pg{again you want to get at why
% these were chosen. Also mention that you searched over a large
% space of possible features and these were the best.}
%
% [EB] Tried addressing. Does it look redundant when read after 'Feature Selection'
% bullet from the previous section?
%
\newline
\newline
\noindent \textbf{Physical Address}: Here we use the lower bits 
of the physical address of 
the demand access that triggers the prefetch. This address corresponds 
to a  stream of accesses that SPP and the PPF have seen before. 
Therefore, PPF will correlate the past behavior of this address 
to the prefetch outcome. 
%
%\pg{what do you mean by ``base address'' the full physical
%  address?  Be specific.  Which bits and why? You should probably also
%  talk about what you do when the number of bits is unwieldy.}
%
\newline
\newline
\noindent \textbf{Cache Line} and \textbf{Page Address}: These two
separate features are derived from the base address that triggered the
prefetch, as follows: $base\_addr >> LOG2\_BLOCK \_SIZE$ and
$base\_addr >> LOG2\_PAGE\_SIZE$ respectively. The idea behind using
three different shifted versions of the same feature is that it allows
the filter to focus its examination in more detail on different
aspects of the address than with a single version.  It also helps give
more importance to the overlapping bits and lesser importance to most
and least significance bits.  This approach can also eliminate
destructive interference that can be caused by directly folding
the address bits into half.
%
%
% djimenez: this is a nice intuition and explanation. 
% [EB] Thank you!
\newline
\newline
\noindent \textbf{Program Counter XOR Depth}: The PC is for the
instruction that triggered the prefetch chain.  Depth refers to the
iteration count of the lookahead stages.  By itself, we find the PC to
not be a good basis for filtering a lookahead prefetcher, as all the
prefetches with depth >= 1 are aliased into the same PC, which will
not be the PC of the eventual actual demand access.  This feature
resolves a PC into a different value for each lookahead depth of
prefetch speculation, giving a more accurate correlation in lookahead
cases.  This is akin to the concept of Virtual Program
Counters~\cite{VPC} introduced by Kim \textit{et. al.} for indirect
branch prediction.
%
\newline
\newline
\noindent \textbf{PC\textsubscript{1} XOR PC\textsubscript{2}>>1 XOR
  PC\textsubscript{3}>>2}: Here $PC_i$ refers to the last $i^{th}$ PC
before the instruction that triggered the current prefetch.  Hashing
together the last three PCs informs PPF about the path that led to the
current demand access and helps capture and branching information of
the current basic block.  PCs are shifted in the increasing order of
history before being hashed together. This is done to avoid the
resultant value of zero when 2 or more PCs are the same. Additionally,
blurring the information as it gets older allows us to get a wider and
yet more approximate look into the program's history.
%
\newline
\newline
\noindent \textbf{Program Counter XOR Delta}: This feature tells us if
a given PC favors particular value(s) of delta.  As noted earlier,
while the PC alone does not convey useful information, this hash
resolves the PC into different values based on the tendency of that PC
to favor a certain delta.  Thus, the dynamic nature of different
instances of the same memory access instruction is captured here.
%
 % GAC: Seems more predictable than dynamic if its capturing 
 % what is essentially a streaming access pattern. Not quite sure how to rephrase,
 % if I'm correct
%
 % GAC: Again having trouble phrasing the below point. I was attempting to say
 % that despite SPP having low confidence due to speculation, PPF may 
 % reinforce prefetches with lower confidence if SPP has had a tendency to be 
 % correct in the past.
\newline
\newline
\noindent \textbf{Confidence}: The confidence, on a scale of 0 to 100,
used to throttle lookahead depth in the original SPP design.  While
the original confidence does not directly make the decision to
prefetch, PPF correlates it to the correctness of a proposed prefetch.
While the original SPP may have dismissed a prefetch due to running
further into speculation, PPF can use the original confidence as
indicator of not only when prefetches become less confident, but also
how likely a low confident speculation is correct in the context of
other features.
%
\newline
\newline
\noindent \textbf{Page Address XOR Confidence}: This feature scores
the tendency of each page to be prefetch friendly or prefetch averse.
It helps resolve a page into different entries depending on its
confidence for prefetching, which can vary during phases of a program
execution.
%
\newline
\newline
\noindent \textbf{Current Signature XOR Delta}: Recall from the
discussion of SPP in Section~\ref{Background-SPP} that the new
signature is generated using the old signature and the current block
delta.  The result of this feature is the next signature that is
predicted to be accessed based on the delta predicted by SPP. While
``Current Signature XOR Delta'' is not the exact formula for
generating the future signature, it gives an approximate idea of the
path that the combination of these two values can lead to.
%
%Gino: Added the sentence below, is this correct?
%[EB]: It gives the next signature instead of next delta
%
%%%%%%% Below are the features rejected when downsizing 14 -> 9
% \item \textbf{Current Signature}: The 12-bit signature used by SPP
%   to index into the Pattern Table
% \item \textbf{Delta}: The signed version of the difference in the
%   block address that triggers the prefetch and the block address of
%   the data being prefetched
% \item \textbf{PC}: Program Counter of the Load instruction that
%   triggers that particular prefetch
% \item \textbf{Base address XOR Delta}: If any particular value(s) of
%   delta are highly favored by a given address which triggered the
%   prefetch, this composite feature can capture that information
% \item \textbf{Cache line XOR Delta}: this feature captures if any
%   particular delta is highly favored by a given cache line address
%\item \textbf{Confidence XOR Delta}: This gives a combined insight
%  into that fact that certain delta-confidence pairs predicted by
%  SPP might correlate more with the prefetch outcome
%%%%%%%%%%%%%%%%%%%%%%
%
\newline
\newline
As can be noted above, some composite features are derived from simple
hashing (XOR) of two primary features. There is always a question of
usefulness of such composite features and the new information added.
We justify the choice of each feature by quantifying the contribution
made towards predicting prefetch behavior, in Section
\ref{Method-Features}.  Finally, as noted above, each feature indexes
into its independent entry of perceptron weights.

% [EB]: Think of more implementation specific details
% 7 to 12 bits of indexing