### 2019-05-22 related.tex
ran spellcheck

\section{Related Work}
\label{related}

\subsection{Spatial Prefetchers}

Spatial prefetchers include such well-understood examples as the
next-line (or next-$n$-line) prefetcher, and the stream prefetcher,
and are distinguished by prefetching data without regard for the order
in which the data will be accessed.  In addition to these simpler
examples, Somogyi \textit{et al.}  propose Spatial Memory Streaming
(SMS)~\cite{SMS}.  SMS works by learning the spatial footprint of all
data used by a program within a region of memory around a given
missing load, and when the load that causes an new miss elsewhere, the
same spatial footprint is prefetched.  Ishii \textit{et al.} propose
the Access Map Pattern Matching prefetcher (AMPM)~\cite{AMPM}, which
creates a map of all accessed lines within a region of memory, and
then analyzes this map on every access to see if any fixed-stride
pattern can be identified and prefetched that is centered on the
current access. DRAM-Aware AMPM (DA-AMPM)~\cite{DA_AMPM} is a variant
of AMPM that delays some prefetches so they can be issued together
with others in the same DRAM row, increasing bandwidth utilization.
Pugsley \textit{et al.}  propose the Sandbox
Prefetcher~\cite{Sandbox}, which analyzes candidate fixed-offset
prefetchers in a sandboxed environment to determine which is most
suitable for the current program phase.  Michaud proposes the
Best-Offset Prefetcher~\cite{BOP}, which determines the optimal offset
to prefetch while considering memory latency and prefetch timeliness.

\subsection{Lookahead Prefetchers}

Unlike spatial prefetchers, lookahead prefetchers take program order
into account when they make predictions.  Shevgoor \textit{et al.}
propose the Variable Length Delta Prefetcher (VLDP)~\cite{VLDP}, which
correlates histories of deltas between cache line accesses within
memory pages with the next delta within that page. SPP~\cite{SPP} and
KPC's prefetching component~\cite{KPC} are more recent examples of
lookahead prefetchers. They try to predict not only what data will be
used in the future, but also the precise order in which the data will
be used, within a given page. Predictions made by lookahead
prefetchers can be fed back into their prediction mechanisms to
predict even further down a speculative path of memory accesses. These
prefetchers can also generalize their learned patterns from one page,
and use those patterns to make predictions in other pages.

\subsection{Managing Prefetched Data}

A low-accuracy aggressive prefetcher can significantly harm
performance.  To minimize interference from prefetching, Wu \textit{et
  al.} propose PACMan~\cite{pacman}, a prefetch-aware cache management
policy. PACMan dedicates some LLC sets to each of three competing
policies that treat demand and prefetch requests differently, using
the policy in the rest of the cache that shows the fewest
misses. Seshadri \textit{et al.} propose ICP~\cite{icp}, which demotes
a prefetch to the lowest reuse priority on a demand hit, based on the
observation that most prefetches are dead after their first hit. To
address prefetcher-caused cache pollution, it also uses a variation of
EAF~\cite{eaf} to track prefetching accuracy, and inserts only
accurate prefetches to the higher priority position in the LRU
stack. Jain \textit{et al.} propose Harmony~\cite{Harmony} to
accommodate prefetches in their MIN algorithm-inspired Hawkeye cache
management system. Ebrahimi \textit{et al.} introduce
HPAC~\cite{HPAC} which provides a coordinated control between 
multiple prefetchers present in a CMP by looking at the 
prefetcher-induced inter-core interference.

\subsection{Machine Learning for Prefetching}

Peled \textit{et al.} introduce interesting ideas for on-line
Reinforcement Learning and dynamically scaling the magnitude of
feedback given to the baseline prefetcher~\cite{Semantics}. The
prefetcher relies on compiler support to receive features and build
the context.  Liao \textit{et al.}  focus on prefetching for data
center applications~\cite{Datacenter}.  They use offline machine
learning algorithms such as SVMs and logistic regression to do a
parametric search for an optimal prefetcher configuration.  Hasheni
\textit{et al.}~\cite{LSTM} categorize prefetching as a regression
problem and use LSTM based Deep Learning approach.

Wang and Lou propose a similar work where perceptrons filter useless
prefetches~\cite{ArxivPerc}.  In their design's primary focus was on
improving the accuracy of an unmodified baseline prefetcher.  Unlike
the scheme presented here, they implement a basic Rosenblatt
perceptron, with general error-correction learning.  While they are
able to increase accuracy, their design results in lower coverage, and
hence has low impact on overall performance.

\subsection{Perceptrons in Cache Management}

In addition to branch prediction~\cite{PerceptronPredictor},
perceptron-based learning has been applied to the area of cache
management.  Teran \textit{et al.} propose using perceptrons to
predict cache line reuse, bypass, and replacement~\cite{Perc_Reuse}.
Perceptron Learning trains weights selected by hashes of multiple
features, including the PC of the memory access instruction, some
other recent PCs, and two different shifts of the tag of the
referenced block. These features are used to index into weight tables,
and the weights are then thresholded to generate a prediction. When a
block from one of a few sampled sets~\cite{sdbp} is reused or evicted,
the corresponding weights are decremented or incremented, according to
the perceptron learning rule. Multiperspective Reuse
Prediction~\cite{Multiperspective} improves on Perceptron Learning by
contributing many new features.